An Incomplete Overview of some Applications of Game Theory to Patient Flow

### 2014-05-13 paper.tex
Finished intro

What damage do the selfish choices of some afflict on to the welfare of all? The work presented in this paper answers this question in the context of public service systems.

There is a substantial quantity of literature on the subject of equilibrium behaviour of a queueing system where congestion is a factor influencing behaviour \cite{Adler1969,Bell1983,Edelson1971,Knudsen1972,Luski1976,Naor1969,Yechiali1972}. This can be traced back to a series of short communications between Leeman \cite{Leeman1964,Leeman1965} and Saaty \cite{Saaty1965}. This paper builds on this literature by considering the problem of public service systems. For most public services, congestion is a negative aspect of service quality. Examples of this are healthcare systems (waiting lists), transports systems (traffic jams) and/or schools (overcrowding of class rooms).

@@

Of course, the motivation for the introduction of choice is to create competition in the hope that this would improve overall service quality. The aim of the work presented here is to use a game theoretical approach to measure the immediate effect of choice (i.e. prior to the effect of competition). The approach proposed is based on a measure called \emph{the price of anarchy} \cite{Chau2003,Cole2006,Correa2008,Gilboa-Freedman2009,Haviv2007,Musacchio2009,Roughgarden2003,Roughgarden2005,Roughgarden2004,Skinner2010}. First introduced in \cite{Koutsoupias1999} (a conference paper that has been reprinted in \cite{Koutsoupias2009}), the price of anarchy is the ratio of the costs of the worst possible Nash equilibrium and the social optimum, and can therefore be interpreted as a measure of the efficiency of a system.

Game theory is the mathematical study of interactive decision making between rational decision makers \cite{Maschler2013}.
A game is a description of strategic interactions that includes the constraints on the actions that the players can take and on the players' interests.

Some other work that has not concentrated on EDs, but has been applied to healthcare includes: \cite{Knight2013}, where results concerning the congestion related implications of decisions made by patients when choosing between healthcare facilities were presented. In \cite{Howard2002} a model of the accept/reject decision for transplant organs is developed.

In the wider intersection of game theory and queueing theory (where this work lies) papers that consider price and/or capacity include: \cite{Allon2007, Cachon2002,  Cachon2007, Kalai1992, Levhari1978}. In these models, the choice of price/or capacity determines the arrival rate for each firm; this is similar to the approach taken in this paper.

\section{Choosing queues}

@@

\section{CCU Work}

@@

\section{Conclusions}

>>>>

Understanding the effect of selfish individuals in queueing systems can be traced back to a series of short communications between Leeman \cite{Leeman1964,Leeman1965} and Saaty \cite{Saaty1965}.

In \cite{Adler1969,Bell1983,Edelson1971,Knudsen1972,Luski1976,Naor1969,Yechiali1972} a variety of models are discussed that obtain optimal and equilibrium behaviour in situations where individuals can choose between a best queue to join.
Most of the work considers these systems in a theoretical context.

@@

This is a question that is answered using the Game theoretical concept referred to as the \textit{Price of Anarchy} (PoA).
The PoA is defined as the ratio of the selfish utility to the optimal utility.

Section \label{sec:choosingqueues} will discuss some PoA analysis applied to individuals choosing between queues.
Section \label{sec:managingqueues} will discuss similar analysis in the case of selfish behaviour related to the management of queues.
Section \label{sec:conclusion} will conclude with a summary and thoughts for further work.

\section{Choosing queues}\label{sec:choosingqueues}

@@

\section{Managing queues}\label{sec:managingqueues}

@@

\section{Conclusions}\label{sec:conclusion}
